---
title: Subjects With Adverse Events by System Organ Class and Preferred Term (Rate and Count, Empty Categories Shown), Safety Population, Pooled Analysis (or Trial X)
subtitle: FDA Table 47
categories: [table, FDA, safety, adverse events]
---

::::: columns

:::: {.column width="52%"}
### FDA IG Shell

<!-- TODO: add ig_shell.png (FDA IG example shell) once available from the source document. -->

### Programming Notes

**Background and Instructions**

This is a companion to Table 46. It summarises treatment-emergent adverse events by system organ class and preferred term, but the system organ class is a factor carrying a category that no subject experiences. Making the expected categories explicit factor levels keeps those rows in the table (rendered as zero) rather than dropping them, so the shell stays stable across studies where some categories may have no events.

**Customization**

Set the factor levels of the grouping variable to the full list of categories the study expects to report.

::::

:::: {.column width="4%"}
::::

:::: {.column width="44%"}
### Output

<!-- TODO: add result.png (screenshot of the rendered table) once available. -->
See the **Table** tab below for the live rendered output.

::::

:::::

::: panel-tabset
## Full Code

```{r fullcode, file="../../../inst/templates/fda-table_47.R", message=FALSE, warning=FALSE, results='hide'}
```

## Table

```{r tbl}
tbl
```

## ARD

```{r ard, message=FALSE, warning=FALSE}
withr::local_options(width = 9999, tibble.print_max = Inf)
print(ard, columns = "all")
```

:::
