---
title: "Getting Started"
toc: true
toc-depth: 4
---

### Environment Setup and Package Installation

#### Environment Setup

Please ensure that R version 3.6 or higher is installed on your computer. This is required to install the `cardinal` package and its dependencies. You can check your current version of R by running `sessionInfo()` within the R console.

#### Installing Package Dependencies

The `cardinal` package requires several packages available on CRAN.

To use the `cardinal` package, ensure you have these necessary package dependencies installed by running the following code:

```{r, echo=TRUE, eval=FALSE}
if (!require("gtsummary")) install.packages("gtsummary")
if (!require("cards")) install.packages("cards")
if (!require("cardx")) install.packages("cardx")
if (!require("formatters")) install.packages("formatters")
if (!require("rtables")) install.packages("rtables")
if (!require("rlistings")) install.packages("rlistings")
if (!require("tern")) install.packages("tern")
```

Optionally, to run the examples provided within the `cardinal` package and on this site you can also install the random CDISC data package, which contains example datasets, by running the following:

```{r, echo=TRUE, eval=FALSE}
if (!require("random.cdisc.data")) install.packages("random.cdisc.data")
```

For more information on these packages, see the [Additional Resources](#additional-resources) section below.

#### Installing the cardinal Package

To install the `cardinal` package, run the following line of code:

```{r, echo=TRUE, eval=FALSE}
if (!require("cardinal")) remotes::install_github("pharmaverse/cardinal")
```

### Loading Data

The table functions provided by `cardinal` are designed to work with any datasets which adhere to CDISC standards. If you have access to CDISC data, load this into your R environment as usual to use when generating tables.

If you do not have access to CDISC data, or want to run the examples provided within the package, you can use the `random.cdisc.data` package to load a selection of datasets containing example synthetic randomized CDISC data. Datasets are stored in the package by name, prefixed by the letter `"c"` (for "cached"). For example, after loading the `random.cdisc.data` package you can access an ADSL dataset by running `cadsl`, or ADAE with `cadae`. This data is loaded in where necessary as part of the provided examples.

### Template Navigation

A list of all templates currently available in `cardinal` is provided in the [Template Library Index](index-templates.qmd), with tables designed based on the [FDA Standard Safety Tables and Figures: Integrated Guide](https://downloads.regulations.gov/FDA-2022-N-1961-0046/attachment_1.pdf).

### Feature Requests

If you would like to request the addition of a new feature to a pre-existing function within `cardinal`, or would like a specific table template to be added, please file an issue on GitHub [here](https://github.com/pharmaverse/cardinal/issues) or reach out to the cardinal team directly via our Slack channel [here](https://app.slack.com/client/T028PB489D3/C04MQS12MND).

You can access the source code for all currently available functions in the `cardinal` GitHub repository [here](https://github.com/pharmaverse/cardinal/tree/main/R).

### Additional Resources {#additional-resources}

See the following packages used by `cardinal` for more information:

-   [`tern`](https://insightsengineering.github.io/tern): Clinical trials analysis functions
-   [`rtables`](https://insightsengineering.github.io/rtables): Table creation
-   [`rlistings`](https://insightsengineering.github.io/rlistings): Listing creation
-   [`formatters`](https://insightsengineering.github.io/formatters): Additional rendering formatting
-   [`random.cdisc.data`](https://insightsengineering.github.io/random.cdisc.data): Example synthetic randomized CDISC datasets (required for examples only)
